A Large-scale Study of Polymorphism in R

### 2019-04-02 notebooks/paper_L123_results_section.Rmd
adjusted l0, looking at l1, matrices next, will update discussion in results to reflect new components of the TS

Step 2: Get data ready for L1.
So: Translate the type system to R.
```{r}
los <- change_type_systems_lopkg(los, type_map_T0_to_r, unparam=T)
```

Step 2.1: Counts.
```{r}
L1_c_basic_df <- build_arg_count_df_from_lopkgs(los)

L1_c_f_basic_df <- build_fun_count_df_from_lopkgs(los)
```

```{r}
xtable(L1_c_basic_df) %>% print(include.rownames=F)
xtable(L1_c_f_basic_df) %>% print(include.rownames=F)
```

Step 2.2: Top.
```{r}
L1_poly_type_args <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los, extract_poly_type_args))
L1_poly_type_args <- L1_poly_type_args[order(-L1_poly_type_args$count),] # sort by count
L1_poly_type_args$perc <- round(L1_poly_type_args$count/sum(L1_poly_type_args$count) * 100, 2)
```

Step 2.3: Print top 10.
```{r}
format_df_for_print(L1_poly_type_args[1:10,]) %>%
xtable %>% print(include.rownames=F)
```

@@

TODO oops scalar vector thing.

@@

Step 5. Translate to L3?
What is L3?
We want class info and attribute info, on some level.
Let's do classes first.
```{r}
# get rid of classes we have dealt with.
los <- lapply(los, function(lof) lapply(lof, class_to_primitive_df))
```

How do we want to deal with the classes?
Should they become a type, if they are not primitive?
Let's look at class poly counts maybe.
```{r}
L3_c_basic_df <- build_arg_count_df_from_lopkgs(los)

@@

Let's find out when class is monomorphic and type is polymorphic.
```{r}
num_poly_type_not_poly_class <- apply_count_to_lopkg_c(los, args_count_all_poly_type_mono_class)
```

Are there patterns?
```{r}
lo_pt_mc <- reduce_extracted_lopkgs_to_one_new(apply_extract_to_lopkg(los, extract_poly_type_args_w_mono_class))
lo_pt_mc <- lo_pt_mc[order(-lo_pt_mc$count),] # sort by count
lo_pt_mc$perc <- round(lo_pt_mc$count/sum(lo_pt_mc$count) * 100, 2)

format_df_for_print_new(lo_pt_mc[1:10,]) -> put_me
put_me
xtable(put_me) %>% print(include.rownames=F)
```

How about attributes?
```{r}
los_a <- lapply(los, function(lofun) lapply(lofun, function(df) convert_df_attr_to_ts(df, type_map_T0_to_r)))

L3_poly_attr_args <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los_a, extract_poly_attr_args))
L3_poly_attr_args <- L3_poly_attr_args[order(-L3_poly_attr_args$count),] # sort by count
L3_poly_attr_args$perc <- round(L3_poly_attr_args$count/sum(L3_poly_attr_args$count) * 100, 2)

>>>>

Step 2: Get data ready for L1.
So: Translate the type system to R.
```{r}
# raw_NA -> NULL
los <- lapply(los, function(lofun) lapply(lofun, make_NAs_into_NULLs))

los <- change_type_systems_lopkg(los, type_map_T0_to_r, unparam=T)

# L <: I <: D
los <- lapply(los, function(lofun) lapply(lofun, logical_int_double_subtype_df))
```

Step 2.1: Counts.
```{r}
L0_c_basic_df <- build_arg_count_df_from_lopkgs(los)

L0_c_f_basic_df <- build_fun_count_df_from_lopkgs(los)
```

```{r}
xtable(L0_c_basic_df) %>% print(include.rownames=F)
xtable(L0_c_f_basic_df) %>% print(include.rownames=F)
```

Step 2.2: Top.
```{r}
L0_poly_type_args <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los, extract_poly_type_args))
L0_poly_type_args <- L0_poly_type_args[order(-L0_poly_type_args$count),] # sort by count
L0_poly_type_args$perc <- round(L0_poly_type_args$count/sum(L0_poly_type_args$count) * 100, 2)
```

Step 2.3: Print top 10.
```{r}
format_df_for_print(L0_poly_type_args[1:10,]) %>%
xtable %>% print(include.rownames=F)
```

@@

@@

Step 5. Translate to L3?
What is L3?
We want class info and attribute info, on some level.
Let's do classes first.
```{r}
# get rid of classes we have dealt with.
los <- lapply(los, function(lof) lapply(lof, class_to_primitive_df))

los_a <- lapply(los, function(lofun) lapply(lofun, function(df) convert_df_attr_to_ts(df, type_map_T0_to_r)))
```

How do we want to deal with the classes?
Should they become a type, if they are not primitive?
Let's look at class poly counts maybe.
```{r}
L3_c_basic_df <- build_arg_count_df_from_lopkgs(los_a)

@@

Let's find out when class is monomorphic and type is polymorphic.
```{r}
# num_poly_type_not_poly_class <- apply_count_to_lopkg_c(los, args_count_all_poly_type_mono_class)
```

Are there patterns?
```{r}
# lo_pt_mc <- reduce_extracted_lopkgs_to_one_new(apply_extract_to_lopkg(los, extract_poly_type_args_w_mono_class))
# lo_pt_mc <- lo_pt_mc[order(-lo_pt_mc$count),] # sort by count
# lo_pt_mc$perc <- round(lo_pt_mc$count/sum(lo_pt_mc$count) * 100, 2)
# 
# format_df_for_print_new(lo_pt_mc[1:10,]) -> put_me
# put_me
# xtable(put_me) %>% print(include.rownames=F)
```

How about attributes?
```{r}
L3_poly_attr_args <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los_a, extract_poly_attr_args))
L3_poly_attr_args <- L3_poly_attr_args[order(-L3_poly_attr_args$count),] # sort by count
L3_poly_attr_args$perc <- round(L3_poly_attr_args$count/sum(L3_poly_attr_args$count) * 100, 2)

### 2019-04-02 paper/main.tex
adjusted l0, looking at l1, matrices next, will update discussion in results to reflect new components of the TS

Figure~\ref{fig:L1argcounts} shows the polymorphicity of argument positions
 in R function.  From there, we build up a picture of the polymorphicity of
 functions as a whole, and the data for this function polymorphism can be
 found in Figure~\ref{fig:L1funcounts}.

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 29 18:10:23 2019
\begin{table}[ht]\label{fig:L1argcounts}\centering\begin{tabular}{lrr}  \hline
Type                   & Count & Percentage \\  \hline
Full Monomorphic       & 784K & 85.76 \\
  Monomorphic in Type & 845K & 92.46 \\
  Monomorphic in Class & 827K & 90.41 \\
  Monomorphic in Attribute Pattern & 825K & 90.27 \\
  Total Seen & 914K & --- \\
   \hline
\end{tabular}
\caption{typeof type system (L0) (arguments)}
\end{table}
% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 29 18:10:23 2019
\begin{table}[ht]
\label{fig:L1funcounts}
\centering
\begin{tabular}{lrr}
  \hline
Type & Count & Percentage \\
  \hline
Full Monomorphic & 144K & 66.54 \\
  Monomorphic in Type & 173K & 80.16 \\
  Monomorphic in Class & 163K & 75.21 \\
  Monomorphic in Attribute Pattern & 161K & 74.37 \\
  Total Seen & 216K & --- \\
   \hline
\end{tabular}
\caption{typeof type system (L0) (functions)}
\end{table}

From Figure~\ref{fig:L1argcounts}, we glean that argument positions aren't
very polymorphic in L0.  An argument which is "Monomorphic in Type" is
monomorphic w.r.t. \code{typeof}, in that it was only ever observed to have
one \code{typeof}.  Similarly, an argument is "Monomorphic in Class" if it
is only ever called with one {\tt class}.  As we said earlier, attributes
are a little more complicated, and an argument which is "Monomorphic in
Attribute Pattern" was only ever observed to have one attribute pattern (and
that pattern may itself contain many attributes).

As for the polymorphism of functions as a whole, we glean from
Figure~\ref{fig:L1funcounts} that a sizeable number of functions are
polymorphic.  Functions are said to be polymorphic if at least one of their
arguments or return are polymorphic.  \AT{Say more?}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 29 18:13:17 2019
\begin{table}[ht]\label{tab:L0toppoly}\centering
\begin{tabular}{lrr}  \hline
type & count & perc \\\hline\\[-3mm]
  \D, \I & 19,799 & 28.73 \\[1mm]
  \D, \l & 5,324 & 7.72 \\ [1mm]
  \D, \sN & 5,152 & 7.48 \\ [1mm]
  \C, \sN & 3,682 & 5.34 \\ [1mm]
  \D, \L & 3,388 & 4.92 \\ [1mm]
  \C, \D & 2,842 & 4.12 \\ [1mm]
  \l, \sN & 2,799 & 4.06 \\ [1mm]
  \C, \l & 1,836 & 2.66 \\ [1mm]
  \C, \I & 1,663 & 2.41 \\ [1mm]
  \C, \I, \D & 1,362 & 1.98 \\[1mm]   \hline
\end{tabular}
\caption{Top polymorphic signatures for the L0 type system.}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:03:30 2019
\begin{table}[ht]
\label{tab:argcountsL1}
\centering
\begin{tabular}{lrr}
  \hline
 Argument Polymorphism & Count & Percentage \\
  \hline
  Full Monomorphic & 781K & 85.37 \\
  Monomorphic in Type & 834K & 91.23 \\
%  Monomorphic in Class & 827K & 90.42 \\
%  Monomorphic in Attribute Pattern & 825K & 90.27 \\
  Total Seen & 914K & --- \\
     \hline
\end{tabular}
\caption{Account of {\it argument} polymorphism in L1.}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:10:46 2019
\begin{table}[ht]
\label{tab:funcountsL1}
\centering
\begin{tabular}{lrr}
  \hline
 Function Polymorphism & Count & Percentage \\
  \hline
  Full Monomorphic & 143K & 65.88 \\
  Monomorphic in Type & 167K & 77.25 \\
%  Monomorphic in Class & 163K & 75.24 \\
%  Monomorphic in Attribute Pattern & 161K & 74.37 \\
  Total Seen & 216K & --- \\
   \hline
\end{tabular}
\caption{Account of {\it function} polymorphism in L1.}
\end{table}

>>>>

Table~\ref{tab:L0argcounts} shows the polymorphicity of argument positions
 in R function.  From there, we build up a picture of the polymorphicity of
 functions as a whole, and the data for this function polymorphism can be
 found in Figure~\ref{tab:L0funcounts}.

\begin{table}[ht]\label{tab:L0argcounts}\centering\begin{tabular}{lrr}  \hline
Type                   & Count & Percentage \\  \hline
%  Full Monomorphic & 785K & 85.87 \\ 
  Monomorphic in Type & 866K & 94.76 \\ 
%  Monomorphic in Class & 827K & 90.41 \\ 
%  Monomorphic in Attribute Pattern & 825K & 90.27 \\ 
  Total Seen & 914K & --- \\ 
   \hline
\end{tabular}
\caption{Argument position polymorphism in L0.}
\end{table}

\begin{table}[ht]
\label{tab:L0funcounts}
\centering
\begin{tabular}{lrr}
  \hline
Type & Count & Percentage \\
  \hline
%  Full Monomorphic & 145K & 66.77 \\ 
  Monomorphic in Type & 186K & 85.78 \\ 
%  Monomorphic in Class & 163K & 75.21 \\
%  Monomorphic in Attribute Pattern & 161K & 74.37 \\
  Total Seen & 216K & --- \\
   \hline
\end{tabular}
\caption{Function polymorphism in L0.}
\end{table}

From Table~\ref{tab:L0argcounts}, we glean that argument positions aren't
very polymorphic in L0.  An argument which is "Monomorphic in Type" is
monomorphic w.r.t. \code{typeof}, in that it was only ever observed to have
one \code{typeof}.  Similarly, an argument is "Monomorphic in Class" if it
is only ever called with one {\tt class}.  As we said earlier, attributes
are a little more complicated, and an argument which is "Monomorphic in
Attribute Pattern" was only ever observed to have one attribute pattern (and
that pattern may itself contain many attributes).

As for the polymorphism of functions as a whole, we glean from
Figure~\ref{tab:L0funcounts} that a sizeable number of functions are
polymorphic.  Functions are said to be polymorphic if at least one of their
arguments or return are polymorphic.  \AT{Say more?}

@@

\begin{table}[ht]\label{tab:L0toppoly}\centering
\begin{tabular}{lrr}  \hline
type & count & perc \\ 
\hline
  \D, \sN & 8301 & 17.33 \\ 
  \D, \l & 6291 & 13.13 \\ 
  \C, \D & 4338 & 9.06 \\ 
  \C, \sN & 4162 & 8.69 \\ 
  \l, \sN & 2912 & 6.08 \\ 
  \C, \l & 1836 & 3.83 \\ 
  \C, \I & 1719 & 3.59 \\ 
  \C, \D, \sN & 1057 & 2.21 \\ 
  \D, error & 906 & 1.89 \\ 
  \I, \sN & 797 & 1.66 \\   \hline
\end{tabular}
\caption{Top polymorphic signatures for the L0 type system.}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:03:30 2019
\begin{table}[ht]
\label{tab:argcountsL1}
\centering
\begin{tabular}{lrr}
  \hline
 Argument Polymorphism & Count & Percentage \\
  \hline
%  Full Monomorphic & 781K & 85.37 \\
  Monomorphic in Type & 834K & 91.23 \\
%  Monomorphic in Class & 827K & 90.42 \\
%  Monomorphic in Attribute Pattern & 825K & 90.27 \\
  Total Seen & 914K & --- \\
     \hline
\end{tabular}
\caption{Account of {\it argument} polymorphism in L1.}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:10:46 2019
\begin{table}[ht]
\label{tab:funcountsL1}
\centering
\begin{tabular}{lrr}
  \hline
 Function Polymorphism & Count & Percentage \\
  \hline
%  Full Monomorphic & 143K & 65.88 \\
  Monomorphic in Type & 167K & 77.25 \\
%  Monomorphic in Class & 163K & 75.24 \\
%  Monomorphic in Attribute Pattern & 161K & 74.37 \\
  Total Seen & 216K & --- \\
   \hline
\end{tabular}
\caption{Account of {\it function} polymorphism in L1.}
\end{table}
